Simple connectome inference from partial correlation statistics in calcium imaging

### 2014-06-19 article.tex
wording comparison state of the art

We will compare our approach to the Pearson correlation (see Section
\ref{sec:inference}), the generalized transfer entropy (GTE) and GENIE3:
\begin{itemize}

\item \textbf{GTE} \citep{stetter2012model}
, the baseline of the challenge, measures for all pairs of
neurons $i$ and $j$ the reduction of uncertainty in predicting future values of
$i$ given the observation of the time series of the neuron $j$
conditionally to the observation of past time series of neuron $i$,
whenever the average fluorescence is over a conditioning level.
Conditioning level was set to reproduce baseline of the challenge.

@@

At first sight, it appears on Fig.\ref{fig:curves} %Tab.\ref{tab:tab1}
that partial correlation outperforms other methods. Two main reasons may explain
this. First, the quality of the infered network seems related to the
conditioning part of the method. In every method except with partial
correlation, indirect links are confounded with direct links.  Secondly, the
gaussian assumption seems verified and thus linear method appears to be
competitive, e.g. the method correlation already gives good results, with a
non-linear method such as GENIE3.

\begin{figure}[bth]
\centering
\includegraphics[width=0.6\linewidth]{images/curves}
\caption{Method precision-recall and roc curves.}
\label{fig:curves}
\end{figure}

Observing Tab.\ref{tab:tab1}, it appears that best methods only infer the
\textbf{undirected} network by giving the same weights to both edges $(i,j)$
and $(j,i)$. This is quite surprising considering the goal of the challenge,
namely predicted directed connections between neurons. Besides, if the true
network undirected, i.e. for all links $(i,j)$, back links are added, is the
infered network then the score obtained according to the ROC AUC metric is
very high ($\approx 0.9956$ on \textit{normal-1}). Therefore, an undirected
infered network seems sufficient to maximize the ROC AUC score without trying
to identify the link direction. Indeed, predicting perfectly directivity only
yields a small gain ($\approx 0.005$ on \textit{normal-1}) which is is not
very rewarding. In contrast, the same experiment with the precision recall
metric rewards less the undirected network ($\approx 0.789$ on
\textit{normal-1}) and the gain margin in predicting the link directivity is
large enough to be interesting ($\approx 0.211$ on \textit{normal-1}). For
these reason, it seems that the metric used in the contest, i.e. the ROC AUC,
does not seem consistent with the goal of predicting directed connections
between neurons. Another metric, such as the AUC PRC, is maybe more
appropriate to assess the quality of the \textbf{directed} infered network.

>>>>

We will compare our approach to the Pearson correlation (see Section
\ref{sec:inference}), the generalized transfer entropy (GTE) and GENIE3.
\begin{itemize}
\item \textbf{GTE} \citep{stetter2012model}
, the baseline of the challenge, measures for all pairs of
neurons $i$ and $j$ the reduction of uncertainty in predicting future values of
$i$ given the observation of the time series of the neuron $j$
conditionally to the observation of past time series of neuron $i$,
whenever the average fluorescence is over a conditioning level.
Conditioning level was set to reproduce baseline of the challenge.
\textcolor{red}{Talk about pre-processing}

@@

Figure \ref{fig:curves} shows the ROC curve and precision-recall curve
for correlation, GTE and partial correlation on the \textcolor{red}{normal-?}
dataset. The partial correlation outperforms other state-of-the-art methods
by at least factor of 2 in AUPRC and an increased of $5\%$ in AUROC. There
are two reasons that explains such overwhelming difference: (i) among
those methods only the partial correlation make the distinction between
direct and indirect connection, (ii) the simulated network from the
\citep{stetter2012model} generator could be reduced to a linear model after
pre-processing.

\begin{figure}[bth]
\centering
\includegraphics[width=0.6\linewidth]{images/curves}
\caption{Method precision-recall and roc curves. \textcolor{red}{GENI3?}}
\label{fig:curves}
\end{figure}

Observing Table \ref{tab:tab1}, it appears that best methods only infer the
\textbf{undirected} network by giving the same weights to both edges $(i,j)$
and $(j,i)$. This is quite surprising considering the goal of the challenge,
namely predicted directed connections between neurons. Besides, if the true
network undirected, i.e. for all links $(i,j)$, back links are added, is the
infered network then the score obtained according to the ROC AUC metric is
very high ($\approx 0.9956$ on \textit{normal-1}). Therefore, an undirected
infered network seems sufficient to maximize the ROC AUC score without trying
to identify the link direction. Indeed, predicting perfectly directivity only
yields a small gain ($\approx 0.005$ on \textit{normal-1}) which is is not
very rewarding. In contrast, the same experiment with the precision recall
metric rewards less the undirected network ($\approx 0.789$ on
\textit{normal-1}) and the gain margin in predicting the link directivity is
large enough to be interesting ($\approx 0.211$ on \textit{normal-1}). For
these reason, it seems that the metric used in the contest, i.e. the ROC AUC,
does not seem consistent with the goal of predicting directed connections
between neurons. Another metric, such as the AUC PRC, is maybe more
appropriate to assess the quality of the \textbf{directed} infered network.
